---
title: "Myopic Loss Aversion in Investment Behavior - Analysis"
author: "Dean K. Young"
date: "January 21, 2016"
output: html_document
---
Please refer to thesis document for commentary.

```{r, echo=FALSE, include=FALSE}
### Initial setup
library(dplyr)
library(magrittr)
library(tidyr)
library(ggplot2)
library(Hmisc)
library(nortest)
library(plm)
library(lme4)
library(sandwich)
library(lmtest)

setwd("C:/Users/Dean/Google Drive/Thesis/Thesis UI and Data/Thesis_Data_Analysis")

data <- read.csv('Exp_Data.csv',header=TRUE)
survey <- read.csv('Exp_Survey.csv',header=TRUE)

survey %<>% mutate(Q3 = ifelse(Q3 == 'Yes',1,ifelse(Q3 == 'No',0,.5))) %>% 
  mutate(Q4 = ifelse(Q4 == 'Yes',1,ifelse(Q4 == 'No',0,.5))) %>%
  mutate(Q5 = ifelse(Q5 == 'Yes',1,ifelse(Q5 == 'No',0,.5)))

survey %<>% select(ID,Q3,Q4,Q5,Q6,Q7)
data %<>% select(-Name,-Student.ID)

wide.data <- inner_join(survey,data,by="ID")
long.data.40 <- wide.data %>% select(ID,Type,num_range("X",161:200)) %>% gather("Choice_Num","Choice",3:42)

remove(data,survey)

```

Type: 1 = Monthly, 2 = Yearly, 3 = Inflated Monthly  
Choices are for Bond allocation (note that the simulator produces choices as for Stock allocation)

### Figure 1
Distribution of final allocation choices for each condition

```{r, echo=FALSE}
hist.data <- wide.data %>% select(Type,Final.Choice) %>% mutate(Final.Choice = Final.Choice * 100) %>%
  mutate(Type = ifelse(Type == 1, "Monthly", ifelse(Type == 2, "Yearly","Inflated Monthly")))

hist.data$Type <- factor(hist.data$Type, levels = c("Monthly", "Yearly", "Inflated Monthly"))

ggplot(hist.data, aes(Final.Choice)) + geom_histogram(binwidth=10, fill="black",color="gray") + 
  facet_wrap(~ Type) +
  scale_x_continuous(breaks=seq(0,100,10)) +
  xlab("Final Allocation to Bond Fund (%)") + 
  ylab("Number of Subjects") 
```

### No figure/table
Kolmogorov-Smirnov (KS) tests

```{r, echo=FALSE}
monthly <- wide.data %>% filter(Type==1) %>% mutate(Final.Choice=Final.Choice*100)
yearly <- wide.data %>% filter(Type==2) %>% mutate(Final.Choice=Final.Choice*100)
infmonthly <- wide.data %>% filter(Type==3) %>% mutate(Final.Choice=Final.Choice*100)

lillie.test(monthly$Final.Choice)
lillie.test(yearly$Final.Choice)
lillie.test(infmonthly$Final.Choice)
```

### Table 1
Percent of portfolio allocated to the Bond Fund for final choice

```{r, echo=FALSE}
ste <- function(x) sqrt(var(x,na.rm=TRUE)/sum(!is.na(x))) #standard error computation

summ.final <- wide.data %>% group_by(Type) %>% 
  summarise(Avg = mean(Final.Choice), sd = sd(Final.Choice), ste = ste(Final.Choice))

t.test1 <- wide.data %>% filter(Type == 1 | Type == 2)
t.test2 <- wide.data %>% filter(Type == 1 | Type == 3)
t.test3 <- wide.data %>% filter(Type == 2 | Type == 3)

t.test(Final.Choice~Type,t.test1)
t.test(Final.Choice~Type,t.test2)
t.test(Final.Choice~Type,t.test3)
```

### Table 2
Percent of portfolio allocated to the Bond Fund for final 40 periods

```{r, echo=FALSE}
summ.40 <- long.data.40 %>% group_by(Type) %>% 
  summarise(Avg = mean(Choice,na.rm=TRUE), sd = sd(Choice,na.rm=TRUE), ste = ste(Choice))

t.2.test1 <- long.data.40 %>% filter(Type == 1 | Type == 2)
t.2.test2 <- long.data.40 %>% filter(Type == 1 | Type == 3)
t.2.test3 <- long.data.40 %>% filter(Type == 2 | Type == 3)

t.test.cluster(t.2.test1$Choice,t.2.test1$ID,t.2.test1$Type)
t.test.cluster(t.2.test2$Choice,t.2.test2$ID,t.2.test2$Type)
t.test.cluster(t.2.test3$Choice,t.2.test3$ID,t.2.test3$Type)
```

### No figure/table
Controlling for survey results

```{r, echo=FALSE}
ctrl.t.test1 <- t.test1 %>% mutate(type.dummy = ifelse(Type==1,0,1), Final.Choice = Final.Choice*100)
ctrl.t.test2 <- t.test2 %>% mutate(type.dummy = ifelse(Type==1,0,1), Final.Choice = Final.Choice*100)
ctrl.t.test3 <- t.test3 %>% mutate(type.dummy = ifelse(Type==2,0,1), Final.Choice = Final.Choice*100)

# Using HC-robust standard errors
coeftest(lm(Final.Choice~type.dummy+Q3+Q4+Q5,ctrl.t.test1), vcov=vcovHC)
coeftest(lm(Final.Choice~type.dummy+Q3+Q4+Q5,ctrl.t.test2), vcov=vcovHC)
coeftest(lm(Final.Choice~type.dummy+Q3+Q4+Q5,ctrl.t.test3), vcov=vcovHC)
```

### Figure 2A-B
Bootstrapped differences in average allocation to the Bond Fund for the monthly and yearly conditions (monthly and inflated-monthly - B)

```{r, echo=FALSE}
bootstrap <- function(a,b){
  set.seed(76)
  boot.samples <- c(rep(NA,100000))
  for (i in 1:100000){
    t.1 <- sample(length(a), length(a), replace=TRUE)
    t.2 <- sample(length(b), length(b), replace=TRUE)
    t <- mean(a[t.1])-mean(b[t.2])
    boot.samples[i] <- t
    
  }
  return(as.data.frame(boot.samples) %>% rename(x=boot.samples))
}

boot.samples.1 <- bootstrap(monthly$Final.Choice,yearly$Final.Choice)
boot.samples.2 <- bootstrap(monthly$Final.Choice,infmonthly$Final.Choice)

CI.1 <- quantile(boot.samples.1$x,c(.005,.995))
CI.2 <- quantile(boot.samples.2$x,c(.005,.995))

boot.samples.1 %<>% mutate(CI = x < CI.1[1] | x > CI.1[2])
boot.samples.2 %<>% mutate(CI = x < CI.2[1] | x > CI.2[2])

ggplot(aes(x, fill=CI), data=boot.samples.1) + geom_histogram(color="gray") + 
  geom_vline(xintercept=CI.1, linetype="dotted", color="red", size= 1) +
  geom_text(label=paste("99% CI: [",round(CI.1[1],2),",",round(CI.1[2],2),"]"), aes(x=mean(boot.samples.1$x),y=12000), size=6) +
  scale_fill_manual(values = c("black", "red")) +
  guides(fill=FALSE) +
  coord_cartesian(xlim=c(-30,70), ylim=c(0,12000)) +
  xlab("Difference in Average Allocation (%)") +
  ylab("Count")

ggplot(aes(x, fill=CI), data=boot.samples.2) + geom_histogram(color="gray") + 
  geom_vline(xintercept=CI.2, linetype="dotted", color="red", size= 1) +
  geom_text(label=paste("99% CI: [",round(CI.2[1],2),",",round(CI.2[2],2),"]"), aes(x=mean(boot.samples.2$x),y=12000), size=6) +
  scale_fill_manual(values = c("black", "red"), name="") +
  guides(fill=FALSE) +
  coord_cartesian(xlim=c(-30,70), ylim=c(0,12000)) +
  xlab("Difference in Average Allocation (%)") +
  ylab("Count")
```

### Figure 3A-C
Monthly (Yearly) (Inflated-Monthly) condition subjects' autocorrelation functions (ACF) of allocation choices

```{r, echo=FALSE}
monthly.data.lm <- monthly %>% select(ID,num_range("X",1:200)) %>%
  gather("Period","Choice",2:201) %>% mutate(Period=as.numeric(substring(Period,2)))

yearly.data.lm <- yearly %>% select(ID,num_range("X",1:200)) %>%
  gather("Period","Choice",2:201) %>% mutate(Period=as.numeric(substring(Period,2))) %>%
  filter(is.na(Choice)==FALSE) %>% mutate(Period = (Period %/% 10) + 1)

infmonthly.data.lm <- infmonthly %>% select(ID,num_range("X",1:200)) %>%
  gather("Period","Choice",2:201) %>% mutate(Period=as.numeric(substring(Period,2)))

ts.monthly <- monthly.data.lm %>% spread(ID,Choice) %>% select(-Period)
ts.yearly <- yearly.data.lm %>% spread(ID,Choice) %>% select(-Period)
ts.infmonthly <- infmonthly.data.lm %>% spread(ID,Choice) %>% select(-Period)


par(mar=c(4,4.5,3,1))
par(mfrow=c(4,5))

for (x in colnames(ts.monthly)){
  acf(ts.monthly[[x]],lag.max=10, plot=FALSE) %>% plot(main = x)
  
}

par(mfrow=c(4,5))

for (x in colnames(ts.yearly)){
  acf(ts.yearly[[x]],lag.max=10, plot=FALSE) %>% plot(main = x)
  
}

par(mfrow=c(4,5))

for (x in colnames(ts.infmonthly)){
  acf(ts.infmonthly[[x]],lag.max=10, plot=FALSE) %>% plot(main = x)
  
}

par(mfrow=c(1,1))
```

### No figure/table
Paired sample t-tests

```{r, echo=FALSE}
paired.data <- long.data.40 %>% group_by(ID) %>% summarise(avg_choice = mean(Choice,na.rm=TRUE)) %>% 
  inner_join(select(wide.data,ID,Type,Final.Choice),by="ID")

t.paired.test1 <- paired.data %>% filter(Type == 1)
t.paired.test2 <- paired.data %>% filter(Type == 2)
t.paired.test3 <- paired.data %>% filter(Type == 3)

t.test(t.paired.test1$avg_choice,t.paired.test1$Final.Choice,paired=TRUE)
t.test(t.paired.test2$avg_choice,t.paired.test2$Final.Choice,paired=TRUE)
t.test(t.paired.test3$avg_choice,t.paired.test3$Final.Choice,paired=TRUE)
```

### Figure 4A-B
Comparing allocations over time (monthly vs. yearly conditions) (vs. inflated-monthly - B)

```{r, echo=FALSE}
monthly.data <- monthly %>% select(num_range("X",1:200)) %>% summarise_each(funs(mean)) %>%
  gather("Period","Avg_Choice",1:200) %>% mutate(Type = "Monthly")

yearly.data <- yearly %>% select(num_range("X",1:200)) %>% summarise_each(funs(mean)) %>%
  gather("Period","Avg_Choice",1:200) %>% filter(is.na(Avg_Choice)==FALSE) %>% mutate(Type = "Yearly")

inf.monthly.data <-  infmonthly %>% select(num_range("X",1:200)) %>% summarise_each(funs(mean)) %>%
  gather("Period","Avg_Choice",1:200) %>% mutate(Type = "Inflated Monthly")

graph.data <- bind_rows(monthly.data,yearly.data) %>% bind_rows(inf.monthly.data)

graph.data %>% filter(Type %in% c("Monthly","Yearly")) %>% 
  ggplot(aes(Period, Avg_Choice, group=Type)) + 
  geom_line(size=2, aes(col=Type)) + 
  scale_x_discrete(breaks=NULL) + 
  scale_y_continuous(breaks=seq(0,100,25)) +
  scale_color_manual(values=c("Blue","Red"),name="Condition") +
  ylab("Average Allocation to Bonds (%)") + xlab("Time (in Periods)") + 
  geom_smooth(method="lm",size=2,color="black",fill=NA)
  
graph.data %>% filter(Type %in% c("Monthly","Inflated Monthly")) %>% 
  ggplot(aes(Period, Avg_Choice, group=Type)) + 
  geom_line(size=2, aes(col=Type)) + 
  scale_x_discrete(breaks=NULL) + 
  scale_y_continuous(breaks=seq(0,100,25)) +
  scale_color_manual(values=c("Red","Blue"), name="Condition",breaks=c("Monthly","Inflated Monthly")) +
  ylab("Average Allocation to Bonds (%)") + xlab("Time (in Periods)") + 
  geom_smooth(method="lm",size=2,color="black",fill=NA)

```

### Table 3
Regressions predicting allocation to Bond Fund (%) from trial number

```{r, echo=FALSE}
### Function to get cluster-robust standard errors
### Source: http://www.ne.su.se/polopoly_fs/1.216115.1426234213!/menu/standard/file/clustering1.pdf
clx <- function(fm, dfcw, cluster){
  M <- length(unique(cluster))
  N <- length(cluster)
  dfc <- (M/(M-1))*((N-1)/(N-fm$rank))
  u <- apply(estfun(fm),2,
             function(x) tapply(x, cluster, sum))
  vcovCL <- dfc*sandwich(fm, meat=crossprod(u)/N)*dfcw
  + coeftest(fm, vcovCL) }

fix.lm1 <- lm(Choice~Period + factor(ID) - 1,monthly.data.lm)


cluster.lm1 <- lm(Choice~Period,monthly.data.lm)


fix.lm2 <- lm(Choice~Period + factor(ID) - 1,yearly.data.lm)


cluster.lm2 <- lm(Choice~Period,yearly.data.lm)

fix.lm3 <- lm(Choice~Period + factor(ID) - 1,infmonthly.data.lm)


cluster.lm3 <- lm(Choice~Period,infmonthly.data.lm)
```

Fixed-Effects (Monthly, Yearly, Inflated-Monthly)
```{r, echo=FALSE}
summary(fix.lm1)
summary(fix.lm2)
summary(fix.lm3)
```

Cluster-Robust SE's (Monthly, Yearly, Inflated-Monthly) 
```{r, echo=FALSE}
clx(cluster.lm1,1,monthly.data.lm$ID)
clx(cluster.lm2,1,yearly.data.lm$ID)
clx(cluster.lm3,1,infmonthly.data.lm$ID)
```

Fixed-Effects and Cluster-Robust SE's (Monthly, Yearly, Inflated-Monthly)
```{r, echo=FALSE}
clx(fix.lm1,1,monthly.data.lm$ID)
clx(fix.lm2,1,yearly.data.lm$ID)
clx(fix.lm3,1,infmonthly.data.lm$ID)
```

### Table 4.
Regressions of final allocation to Bond Fund (%) on various measurements of finance knowledge and recollection of stock value loss.

```{r, echo=FALSE}
sum(wide.data$Q6 != "DK")
sum(wide.data$Q7 != "DK")

survey.lm1 <- wide.data %>% select(Final.Choice,Q6,Q7) %>%
  mutate(tech.dummy=ifelse(Q6 == "DK",0,1),fin.dummy=ifelse(Q7 == "DK",0,1),Final.Choice=Final.Choice*100) 
survey.lm2 <- survey.lm1 %>% filter(Q6 != "DK") %>% mutate(Q6=as.numeric(as.character(Q6))*100)
survey.lm3 <- survey.lm1 %>% filter(Q7 != "DK") %>% mutate(Q7=as.numeric(as.character(Q7))*100)

summary(lm(Final.Choice~tech.dummy + fin.dummy,survey.lm1))
summary(lm(Final.Choice~Q6,survey.lm2))
summary(lm(Final.Choice~Q7,survey.lm3))
```