---
title: "Myopic Loss Aversion in Investment Behavior - Appendix B Simulation"
author: "Dean K. Young"
date: "February 4, 2016"
output: html_document
---

```{r, include=FALSE}
library(psych)
library(truncnorm)
library(plyr)
library(dplyr)
library(ggplot2)
library(magrittr)
library(nortest)
```

### Simulation Code (Not set to run)
This code creates the "dist_data.csv" file that contains the distributions of the simulated data.

```{r, eval=FALSE}
dist <- rep(NA,6*100000)
n <- rep(NA,6*100000)
base <- 0
set.seed(76)

ptm <- proc.time() # Start timer

for (i in c(10,30,50,100,200,400)){ 
  
  for(j in 1:100000){
    a <- rnorm(i,1.01,.0354)
    b <- rtruncnorm(i,a=1,b=Inf,mean=1.0025,sd=.00177)
    diff <- geometric.mean(a) - geometric.mean(b)
    dist[j+(100000*base)] <- diff
    n[j+(100000*base)] <- i
  }
  base <- base + 1
}

proc.time() - ptm # Check computing time


data <- bind_cols(as.data.frame(dist),as.data.frame(n))
write.csv(data, file="dist_data.csv",row.names=FALSE)
```

### Figure 5.
Simulated distribution of difference in geometric mean return between the Stock Fund and the Bond Fund over various sample sizes

```{r}
data <- read.csv("dist_data.csv")

data %<>% mutate(dist = dist* 100,
                 perform = ifelse(dist < 0,"Underperform","Outperform"),
                 perform_count = ifelse(dist < 0,1,0))

data$n <- factor(data$n, levels=c(10,30,50,100,200,400), labels=c("n = 10", "n = 30", "n = 50", "n = 100", "n = 200", "n = 400"))

data.cor <- ddply(.data=data, .(n), summarize, lab=paste("Prob. of Underperform: ", sum(perform_count)/1000, "%", sep=""))

ggplot(data,aes(dist, fill=perform)) + geom_histogram(color="gray") +
  xlab("Difference (%)") + ylab("Count") + 
  scale_fill_manual(values = c("black", "red"), name="") +
  facet_wrap(~ n) +
  geom_text(data=data.cor, aes(x=.66, y=51000, label=lab), color="red", inherit.aes=FALSE, parse=FALSE)
```

### No figure/table
Distribution means and KS-tests on distributions

```{r}
mean(data$dist)
data %>% group_by(n) %>% summarise(mean(dist))

ks.test <- data %>% filter(n == "n = 10") %>% select(dist)

for (x in c("n = 30", "n = 50", "n = 100", "n = 200", "n = 400")){
  ks.test <- bind_cols(ks.test, data %>% filter(n == x) %>% select(dist))
  
}

colnames(ks.test) <- c(10,30,50,100,200,400)

lillie.test(ks.test$"10")
lillie.test(ks.test$"30")
lillie.test(ks.test$"50")
lillie.test(ks.test$"100")
lillie.test(ks.test$"200")
lillie.test(ks.test$"400")
```

### Figure 5 (Partial)
Comparison of underperformance: Probs for assuming continuously compounded return

```{r}
mu.r <- .0075
sigma2.a.b <- .0354^2 + .00177^2
n.list <- c(10,30,50,100,200,400)
table <- matrix(nrow=1,ncol=2)
for (n in n.list){
  prob <- pnorm(0,mean=mu.r,sd=sqrt(sigma2.a.b/n))
  table <- rbind(table,c(n, paste(round(prob*100,3),"%")))
  
}
```
